---
title: "AABB Results"
author: "Jenan El-Hifnawi"
format: 
 html: 
    title-block-banner: true
    theme: pulse
editor: visual
echo: false
---

*This document outlines the data analysis process for [Ask A Bumble Bee.](https://u.osu.edu/askabumble/) It includes various tables, figures, and the code chunks involved in the creation of these objects. If you have any questions, please contact bumblebeecount\@gmail.com .*

```{r}
#| label: load_packages_and_functions
#| warning: false

if (!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse,
               patchwork,
               here,
               taxize,
               flextable,
               cowplot,
               writexl,
               viridis,
               update = FALSE)

if(!exists("Generalharmonthreattmodeltable", mode = "function"))
  source(here::here("functions",
                    "Generalharmonthreattmodeltable.r"))


if (!exists("generalpilotanalysistable", mode = "function"))
  source(here::here("functions",
                    "GeneralPilotAnalysisTable.R"))

```

```{r}
#| label: load_cleanedish_data_files
#| warning: false

genusleveldatafileinprogress <-
  read_csv(here("data",
                "cleaned.data",
                "FinalGenusLevelAnalysisFileReRankedFlRank.csv"),
           name_repair = "universal")

genusleveldatafileinprogressunfiltered <- 
  read_csv(here("data",
               "cleaned.data",
               "FinalGenusLevelDatafileUnfiltered.csv"))

 rawsurveyconditionstable <-
  read_csv(here("data",
                "raw.data",
                "SurveyConditionsTableBetterDates.csv"),
           name_repair = "universal")
 
 BeeTable <- #7300 rows 
  read_csv(here("data",
                "raw.data",
                "BeeTable29SEP23.csv"),
           name_repair = "universal")
 
 genustofamilytable <-
  read_csv(here("data",
                "cleaned.data",
                "genustofamilytable.csv"),
           name_repair = "universal") %>%
  mutate(FGenus = paste0(Genus," species"))
 
 FlowerTable <- 
   read_csv(here("data",
                  "raw.data",
                  "FINALFLOWERTABLE_GENUS.csv"),
             name_repair = "universal")
 
 FlowerTableAllDupesPresent <- 
      read_csv(here("data",
                  "raw.data",
                  "FINALFLOWERTABLE_GENUS.csv"),
             name_repair = "universal")
```

# **Universal Data Wrangling**

Here, I'm merging in my surveys conditions data, adding in a bee score field, adding family IDs, and filtering out surveys that should be universally thrown out! Surveys removed here include those which had under 5 flower genera, under 5 bees, or the errorcode "NoFRHASBEES". This error code indicates that at least one plant in the survey which was visited by a bee was missing a floral rank value, making the survey unusable. I save the final nice table out at the end of this chunk.

### Universal merging, mutating, and filtering

```{r}
#| label: universal_data_wrangling_bee_table_prep
#| warning: false
#| echo: false
#| code-fold: true

BeeTablePrep1 <- 
  BeeTable %>%
  pivot_wider(.,
              names_from = BSpecies,
              values_from = Count,
              values_fill = 0,
              names_repair = "universal") %>%
  select(!c(Sex, ID, Xylocopa.micans, Bombus.borealis, Bombus.ternarius,Bombus.vagans)) %>%
  left_join(FlowerTable, by = c("SurveyNum", "FNum")) %>% 
  select(!c("IDnew","CountofLots":"corrFloralRank")) 

```

```{r}
#| label: clearing_out_duplicated_rows_MUST_RUN_THIS_CHUNK
#| warning: false
#| echo: false
#| code-fold: true

#In this chunk I adress duplication within the bee table by summing all columns for any rows with the same FGenus and SurveyNum. This needs to occur because each different sex or species will have its own row and that is not needed now. Additionally, flower duplication can cause this duplication. 
# I do so by mapping through a function which groups by both SurveyNum and FNum and calculates the sum, in order to 


variable_names <- names(dplyr::select(BeeTablePrep1,
                                      Bombus.impatiens:Bombus.terricola))

# Function to sum bees on plants of the same genus within a survey
sumbeetabledupes <- function(variable) {
  BeeTablePrep1 <<- BeeTablePrep1 %>%
    group_by(SurveyNum, FGenus) %>%
    mutate(!!variable := sum(!!sym(variable))) %>%
    ungroup()
}

# Use map to apply the function to each variable
map(variable_names, ~sumbeetabledupes(.x))


CleanWideBeeTable <- 
  BeeTablePrep1 %>%
  select(!FNum) %>%
  distinct() %>%
  group_by(SurveyNum) %>%
  rowwise() %>% #makes the following sum operations occur within a row across multiple columns
  mutate(AllBombus = sum(c_across(matches("Bombus")), na.rm = TRUE),
         AllXylocopa = sum(c_across(matches("Xylocopa")), na.rm = TRUE)) %>%
  ungroup() %>%
  relocate(c(FGenus, AllBombus, AllXylocopa), .before = Bombus.impatiens)

```

```{r}
#| label: merging_bee_data_with_genusleveldatafileinprogress
#| warning: false
#| echo: false
#| code-fold: true

##OH MY GOD THIS WORKS!!! PRECIOUS!! The math seems good given that when the BBcount >5 filter is applied there are 11542 rows, as we expect! Once I take that filter away given that it is not relevant to us, I get 13810 rows! 

cleanedgenusleveldataset <- 
  genusleveldatafileinprogressunfiltered %>%
  left_join(genustofamilytable, by = "FGenus") %>%
  left_join(CleanWideBeeTable, by = c("SurveyNum", "FGenus")) %>%
  mutate(across(AllBombus:Bombus.terricola, ~replace(., is.na(.), 0))) %>%
  group_by(SurveyNum) %>% #in the next line I sum count on duplicated genera
  mutate(across(AllBombus:Bombus.terricola, ~ sum(.), .names = "sum_{.col}")) %>%
  ungroup() %>%
  left_join(rawsurveyconditionstable, by = "SurveyNum") %>%
  dplyr::group_by(SurveyNum) %>%
  dplyr::filter(sum(BBcount) > 4) %>% #Remove surveys with under 5 bees
  #dplyr::mutate("SumofBBcount" = sum(BBcount)) %>% #to verify that the filter worked
  dplyr::mutate("NumberofFlowers" = n()) %>% #to verify that the filter worked
  dplyr::filter(NumberofFlowers > 4) %>% #survey 88 should no longer be present
  dplyr::ungroup() %>%
  relocate(ID:Longitude, .before = SurveyNum) %>%
  filter(!grepl("NoFRHASBEES", ErrorCode)) %>% # removes surveys with that error code
  dplyr::add_count(FGenus, name = "FlowerSampleSize")  %>%
  dplyr::mutate("corrFloralRank" = CorrRankAvgFlRank) %>%
  dplyr::mutate("FloralRank" = RankAvgFloralRank) %>%
  dplyr::select(!c(CorrRankAvgFlRank, TestCorrFlRank, RankAvgFloralRank))

```

#### Fowler-adapted Harmon-Threatt table

This table is based on Jarrod Fowler's adaptation of a preference model from [Harmon-Threatt et. al. 2016](doi:10.5061/dryad.6kn05)

```{r}
#| label: tbl-harmon_threatt_model
#| tbl-cap: Jarrod Fowler's adaptation of a preference model from Harmon-Threatt et. al. 2016. This table includes all genera present on 30 or more surveys, sorted in descending order by the final "preference" value. 

#Set the counters

dat <- cleanedgenusleveldataset
cont.y.var <- names(dplyr::select(cleanedgenusleveldataset,
                                        AllBombus:Bombus.terricola))

###CHECK IF I NEED THIS FIRST LOOP - IT SEEMS TO BE DUPLICATED NEEDLESSLY 
# a loop to produce the filtered datasets
for (i in seq_along(cont.y.var)) {

  # Create the file and assign them to objects named such that we can retrieve all of the ones we want later.

  assign(
    paste0(cont.y.var[i],".filtered_data"),

    # Generate the contents of the objects using our function.
    {
      # Extract the variable name without the "sum_" prefix
      var_name <- sub("^sum_", "", cont.y.var[i])

      # Filter the dataset based on the condition
      filtered_dat <- dat %>%
        filter(get(paste0("sum_", var_name)) > 0)

      # Return the filtered dataset
      filtered_dat
    }
  )
}


# a loop to produce the final harmon threatt model tables
for (i in seq_along(cont.y.var)) {
  
  # Create the file and assign them to objects named such that we can retrieve all of the ones we want later.
  
  assign(
    paste0("harmonthreatt.", cont.y.var[i]), 
    
    # Generate the contents of the objects using our function.
    {
      # Extract the variable name without the "sum_" prefix
      var_name <- sub("^sum_", "", cont.y.var[i])
      
      # Filter the dataset based on the condition
      filtered_dat <- dat %>%
        filter(get(paste0("sum_", var_name)) > 0)  
      
      # Use the filtered dataset in the analysis
      harmonthreattmodeltable(filtered_dat, cont.y.var[i],"FGenus")
    }
  )
}
```

```{r}
#| label: saving_harmon_threatt_tables_and_respective_datasets


# Get a list of all objects in the environment
all_objects <- ls()

# Filter objects that begin with the string "harmonthreatt."
harmonthreatttables <- grep("^harmonthreatt\\.", all_objects, value = TRUE)

# Create a list of the selected objects
harmonthreatttables_list <- mget(harmonthreatttables)

# save(
#   harmonthreatttables_list,
#   file = here::here("data",
#               "cleaned.data",
#               "beespeciesharmonthreatttables.RData"))


# Write the list of data frames to an Excel file

 # writexl::write_xlsx(harmonthreatttables_list,
 #                    here("data",
 #                        "summarized.data",
 #                        "genuslevelharmonthreatttables.allbeespecies.samplesize10.xlsx"))


# Filter objects that end with the string "filtered_data"
beespeciesdatasubsets <- grep("filtered_data", all_objects, value = TRUE)

# Create a list of the selected objects
beespeciesdatasubsets_list <- mget(beespeciesdatasubsets)


# save(
#   beespeciesdatasubsets_list,
#   file = here::here("data",
#               "cleaned.data",
#               "beespeciesdatasubsets.RData"))

# writexl::write_xlsx(beespeciesdatasubsets_list,
#                    here("data",
#                        "cleaned.data",
#                        "beespeciesdatasubsets.xlsx"))


```

#### Pilot Analysis Table

```{r}
#| label: PILOT ANALYSIS: DATAFRAME PREP 

# Specify the additional columns you want to select
additional_columns  <- names(dplyr::select(cleanedgenusleveldataset,
                                           ID:Family))

######################

# Iterate through the list and extract the relevant species column
result <- lapply(names(beespeciesdatasubsets_list), function(df_name) {
  # Extract the column name without the ".filtered_data" suffix
  column_name <- sub("\\.filtered_data$", "", df_name)
  
  # concatonate your additional column and species columns
  selected_columns <- c(additional_columns, column_name)
  
  # Extract the specified columns
  selected_data <- beespeciesdatasubsets_list[[df_name]][, selected_columns, drop = FALSE]
  
  return(selected_data)
})

######################

#renaming the dataframes within the list
colnames <- lapply(names(beespeciesdatasubsets_list), function(df_name) {
  # Extract the column name without the ".filtered_data" suffix
  column_name <- sub("\\.filtered_data$", "", df_name)
  return(column_name)
})

renamed_filtered_dataframes <- setNames(result, colnames)

######################

# Create and apply a function to rename the 23rd column to "Count"

# Create the function 
rename_column <- function(dataset) {
  colnames(dataset)[20] <- "Count"
  return(dataset)
}

# Apply the function to each dataset in the list
dataframesforpilotanalysis <- lapply(renamed_filtered_dataframes, rename_column)

######################
# Create the SumofCount, PropBees, and BScore columns! 

finaldataframesforpilotanalysis <- map(
  dataframesforpilotanalysis,
  function(df) df %>%
    group_by(SurveyNum) %>%
    mutate(SumofCount = sum(Count)) %>%
    ungroup() %>%
    mutate(PropBees = Count /SumofCount) %>%
    mutate(BeeScore = PropBees * corrFloralRank) 
)

# save(
#   finaldataframesforpilotanalysis,
#   file = here::here("data",
#               "cleaned.data",
#               "dataframesforpilotanalysistables.RData"))

#writexl::write_xlsx(finaldataframesforpilotanalysis,
#                   here("data",
#                       "cleaned.data", 
#                       "datasetsforpilotanalysis.xlsx")) 

# Plug in the list of final datasheets into the pilot analysis table

finalpilotanalysistables <- map(finaldataframesforpilotanalysis,
                              ~generalpilotanalysistable(., "FGenus"))

# save(
#   finalpilotanalysistables,
#   file = here::here("data",
#               "cleaned.data",
#               "finalpilotanalysistables.RData"))


# writexl::write_xlsx(finalpilotanalysistables,
#                   here("data",
#                       "summarized.data",
#                       "genuslevelpilotanalysistables.allbeespecies.xlsx"))
```

@tbl-pilot_analysis explained:

*NumberSurveys* - The number of surveys that a flower genus appeared in.

*Avg.Corr.RankAvg.FloralRank* - Average corrected rank-average of floral rank. Here we took the rank-average of floral rank on a per survey basis, and then "corrected" that value by scaling it between 0 and 10 with the following calculation:$$CorrRankAvgFlRank = RankAvgFloralRank * 10/(max(RankAvgFloralRank))$$

The value displayed here is the average of this value across all surveys.

*AvgBBCount* - Average bumble bee count - Total number of bumble bees found on a flower genus across all surveys, divided by the total number of surveys that the flower genus appeared in.

*AvgCorrBBCount* - Average corrected bumble bee count. The number of bumble bees on a flower genus in one survey, divided by the total number of bumble bees in that survey. The value displayed here is the average of this value across all surveys.

*BBScore.CorrRankAvgFlRankXCorrBBCount* - Average of Adjusted bumble bee count times adjusted floral rank - Bumble bee count multiplied by adjusted floral rank on a per survey basis. This value accounts for both floral abundance and bee visitation, with highest numbers representing plants with lots of bees despite low abundance and lowest numbers representing plants with few visitations despite high abundance.

#### Combined Summary Table

The "preference" metrics include:

a)  ProportionSurveysWithVisitation: Number of surveys with at least one bumble bee visit) / (total number of surveys with the plant genus

b)  BBScore.CorrRankAvgFlRankXCorrBBCount - Derived from Sam and Jenan's pilot analysis method

c)  JarrodModel.BBScore - vRank derived from Fowler adapted Harmon-Threatt model.

```{r}
#| label: tbl-combined_summary
#| warning: false
#| code-fold: true
#| tbl-cap: A compilation of various "preference" metrics and a couple key values involved in their creation. 

library(readxl)
unfilteredBombusAndXyloCombinedSummaryTables <- read_xlsx(here("data",
                                             "summarized.data", 
                                             "Test.UpdatesToHarmonThreattModel.BombusAndXylocopaGenusLevelSummaryTables.xlsx")) 

test.filteredcolumnsforcomparison <- finalpilotanalysistables[["AllBombus"]] %>%
  select(c(FGenus, NumberSurveys, AvgBeeScore, BulkBeeScore)) %>%
  rename_with(~paste0(., ".filt"))


combinedsummarytable.AllBombus <-
  unfilteredBombusAndXyloCombinedSummaryTables %>%
  dplyr::left_join(test.filteredcolumnsforcomparison, by = c("FGenus" = "FGenus.filt"))

writexl::write_xlsx(combinedsummarytable.AllBombus,
                    here("data",
                        "summarized.data", 
                        "BombusGenusLevelSummaryTable.Testof5orMoreBeeFilter.xlsx")) 


combinedsummarytable.AllXylocopa <-
finalpilotanalysistables[["AllXylocopa"]] %>%
  dplyr::left_join(harmonthreatt.AllXylocopa, by = "FGenus") %>%
  dplyr::select(c(FGenus:BulkBeeScore, vRank)) %>% #remove all the other Harmon-Threatt columns except vRank)
  dplyr::rename("NumberSurveys" = "NumberSurveys.x") %>%
  dplyr::rename("Fowler.HarmonThreattModel.vRank" = "vRank")

```

```{r}
#| label: saving_summary_tables_out
#| warning: false
#| code-fold: true

combined.summary.tables.list <-
  list(
    "combinedsummarytable.AllBombus" = combinedsummarytable.AllBombus,
    "combinedsummarytable.AllXylocopa" = combinedsummarytable.AllXylocopa
  )

# writexl::write_xlsx(combined.summary.tables.list,
#                    here("data",
#                        "summarized.data", 
#                        "BombusAndXylocopaGenusLevelSummaryTables.xlsx")) 


```
